BIP-Predictor

### 2025-05-09 Sweave/BIP_predictor.Rnw
up

# Note: too complex designs (too many indicators) lead to overfitting and thus worse out-of-sample performances
# To illustrate the direct predictor consider the following example of a h-step ahead direct forecast:
f_vec_unfiltered<-NULL
for (h in 0:3)
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
  # Regress selected indicators on forward-shifted BIP
  lm_obj<-lm(forward_shifted_BIP~x_mat_wc[,select_direct_indicator])
  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
  summary(lm_obj)
  sum_obj<-summary(lm_obj)
  f_vec_unfiltered<-c(f_vec_unfiltered,sum_obj$fstatistic["value"])

@@

par(mfrow=c(2,2))
f_vec_hp<-t_mat_hp<-NULL
for (h in 0:3)
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
  # Regress selected indicators on forward-shifted BIP
  length(forward_shifted_BIP)
  lm_obj<-lm(forward_shifted_BIP~t(hp_c_array_wc[,,h+1]))
  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
  sum_obj<-summary(lm_obj)
  sum_obj<-summary(lm_obj)
  f_vec_hp<-c(f_vec_hp,sum_obj$fstatistic["value"])
  t_mat_hp<-rbind(t_mat_hp,sum_obj$coef[-1,4])
  # Technical note: 
  # -Residuals are subject to heteroscedasticity (crises) and autocorrelation
  # -Therefore classic OLS tests for statistical significance are biased
  # -We shall rely on HAC-adjusted p-values further down (R-package sandwich)

@@

mat_re<-rbind(f_vec_unfiltered,f_vec_hp)
rownames(mat_re)<-c("Direct forecast","Direct HP forecast")
colnames(mat_re)<-paste("h=",0:3,sep="")
mat1<-round(mat_re,3)
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat_re)[2]+1),
paste("F-statistics of direct forecasts and direct HP forecasts for forecast horizons $h=0,1,2,3$: fully in-sample.  "),
label=paste("f_stat",sep=""),
center = "centering", file = "", floating = FALSE)
@
We now propose a more refined causal filter design aiming for statistical significance of the resulting predictor at longer forecast horizons, out-of-sample.

>>>>

# Note: too complex designs (too many indicators) lead to overfitting and thus worse out-of-sample performances
# To illustrate the direct predictor consider the following example of a h-step ahead direct forecast:
f_vec_unfiltered<-p_val_unfiltered<-NULL
for (h in 0:3)
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
  # Regress selected indicators on forward-shifted BIP
  lm_obj<-lm(forward_shifted_BIP~x_mat_wc[,select_direct_indicator])
  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
  summary(lm_obj)
  sum_obj<-summary(lm_obj)
  f_vec_unfiltered<-c(f_vec_unfiltered,sum_obj$fstatistic["value"])  
  p_val_unfiltered<-c(p_val_unfiltered,pf(sum_obj$fstatistic["value"], sum_obj$df[1],sum_obj$df[2], lower.tail = FALSE))

@@

par(mfrow=c(2,2))
f_vec_hp<-t_mat_hp<-p_val_hp<-NULL
for (h in 0:3)
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
  # Regress selected filtered indicators on forward-shifted BIP
  length(forward_shifted_BIP)
  lm_obj<-lm(forward_shifted_BIP~t(hp_c_array_wc[,,1]))
  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
  sum_obj<-summary(lm_obj)
  sum_obj<-summary(lm_obj)

  p_val_hp<-c(p_val_hp,pf(sum_obj$fstatistic["value"], sum_obj$df[1],sum_obj$df[2], lower.tail = FALSE))
  f_vec_hp<-c(f_vec_hp,sum_obj$fstatistic["value"])
  t_mat_hp<-rbind(t_mat_hp,sum_obj$coef[-1,4])
  # Technical note: 
  # -Residuals are subject to heteroscedasticity (crises) and autocorrelation
  # -Therefore classic OLS tests for statistical significance are biased
  # -We shall rely on HAC-adjusted p-values further down (R-package sandwich)

@@

mat_re<-rbind(f_vec_unfiltered,f_vec_hp)
mat_re<-rbind(p_val_unfiltered,p_val_hp)
rownames(mat_re)<-c("Direct forecast","Direct HP forecast")
colnames(mat_re)<-paste("h=",0:3,sep="")
mat1<-round(mat_re,5)
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat_re)[2]+1),
paste("P-values of of F-statistics of direct forecasts and direct HP forecasts for forecast horizons $h=0,1,2,3$: fully in-sample.  "),
label=paste("f_stat",sep=""),
center = "centering", file = "", floating = FALSE)
@
We now propose a more refined causal filter design aiming for statistical significance of the resulting predictor at longer forecast horizons, out-of-sample.
